The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-26 paper.tex
sections 1, 2, 3 are so far complete (first robust draft)

We openly explored long-term interactions between human users and a domestic
service robot in a 6-months lasting ethnographic study~\citep{fink_living_2013}. 
We gave a commercially
available vacuum cleaning robot (iRobot's Roomba, see Figure
\ref{fig:aibo-roomba-ipad} middle) to nine households. 
%in the French speaking
%part of Switzerland. 
%The fairly diverse sample of households consisted of 2
%single homes, 1 couple, and 6 families with children in their home. In total,
%this summed up to 30 participants: 15 adults (26-71 years old, mean age 43.6
%years), and 15 children (6 months to 18 years old). During the 6 months, each
%household was visited 5 times to conduct interviews and make field observations
%in the households. Moreover, during several weeks, households were asked to fill
%in a diary to report the usage of the robot and other cleaning activities.
Hereafter, we refer to this study as the ``Roomba study''. It is one of very few 
long-term human-robot interaction studies in the domestic environment. 
In the following, we present findings in respect of two aspects: First, we
qualitatively describe what kinds of anthropomorphic interactions between
participants and the robot were observed and outline how these interactions
evolved over time.  Second, we present people's perception of human-likeness and
social capabilities of the robot and how this perception changed over
time.\footnote{\hl{note that these two different kinds of data
(behavior/interaction and perception) already show that we try to measure
anthropomorphism from these two angles!}}

@@

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize the artifact / 
agent. This perspective assumes that humans directly respond to life-like and social 
cues that the non-human agent or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it, and in turn apply 
human-human social schemas and norms to the occurring interactions.
This view is acceptable since humans naturally react to specific 
cues. From early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
or to respond to social cues, such as facial expressions, social gestures or turn-taking. 
One has to keep in mind that these cues are of various senses and modalities in human-
human interaction, and consequently human-like design also addresses several senses. 
One of the most outstanding aspects is probably the outer appearance of an agent or 
artifact (in particular when it is unknown or unfamiliar). \cite{schmitz_concepts_2011} 
argues that an agent's outer appearance is likely to have an important impact on how 
the object is overall perceived. Consequently, if a robot's outer form resembles a 
human, the robot is likely to be treated similar to a human \hl{(\eg eye-contact is 
established or the robot is greeted)}.

@@

\paragraph*{Anthropomorphic form can be advantageous for HRI\\}

@@

\paragraph*{Anthropomorphic form can be disadvantageous for HRI\\}

@@

Literature on anthropomorphism is quite diverse. There exist different understandings of what 
anthropomorphism is. 
One reason for this might be that anthropomorphism
is a concept that is studied in very different domains that might each
integrate diverse or even contradictory understandings, or simply refer to
different things.
However, the conflict is that, despite the fact that there is no commonly accepted
definition, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed upon
\citep{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002} and \cite{epley_when_2008}
argue that these terms might even be misused. For instance, some researchers refer to \textit{"the
robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others
disagree to such a statement because they argue that the system or artifact itself does not \textit{"contain anthropomorphism" per se} but only gives rise to the process of anthropomorphizing in a
given user and situation. Consequently, \cite{persson_anthropomorphism_2000} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user, and has to be understood as such. 
We will come back to this view later, in the summary of this section (Subsection \ref{sec:multi-factors}).

Paradoxically, while human's tendency to anthropomorphize is commonly observed it is 
at the same time still rather poorly understood
\cite{epley_seeing_2007}. The phenomenon itself has been found to be very complex though sometimes
subtle and as such hard to study.
Also, the methods and measurements applied to study anthropomorphism appear insufficient, when applying a holistic understanding of anthropomorphism emerging from an interaction between a human user and a robot, for instance. Our point here is that to study a phenomenon as complex as anthropomorphism, we believe that one needs to take into account many factors (see Section \ref{sec:multi-factors}), besides looking at the human-like design of the artifact.
However, most of the
experiments that investigate anthropomorphism in robotics are conducted in
controlled lab-settings during short-term human-robot interactions. This
experimental setting can be critical for studying
anthropomorphism and findings might possibly lead to interpretations that would not be generalizable to natural settings or real interaction scenarios.\footnote{For instance, some studies are about showing pictures of different robots to human subjects and then asking them how "human-like" they would rate these robots. What points of reference do the participants have in this setting? What can you really tell from a picture? I found that some normal people totally misinterpret the size / height of a robot, when it is not indicated. Also, when showing only pictures, movements, behaviors, materials, and so forth are totally disregarded. However, these aspects are also crucial to assess how "human-like" a robot appears. Last but not least, it has been found that "human-likeness" is a quite diverse and abstract term that includes many things. Instead of asking such a big question, \cite{kahn_jr._robotic_2006} suggest to ask for more concrete constructs that are typical or unique of the concept of "human-likeness", such as being hungry or feeling guilty. For instance, \cite{ruijten_introducing_2014} propose a 25 concrete questions to measure anthropomorphism along a scale that accounts for lower and higher levels of it. \hl{(let's move this to the measurement section at the end.)}}
Some studies (e.g.
exploring characteristics of the "uncanny valley") also seem to contrast two or
more different anthropomorphic or mechanical systems that might not be
appropriate for comparison given the complex context. Others might also use
empirical methods inappropriate for the given research question which leads to
inaccurate results. \hl{say more}

Anthropomorphism is used in different senses throughout different disciplines 
\citep{duffy_anthropomorphism_2003} which makes it difficult to synthesize findings. 
Depending on which understanding of anthropomorphism is applied, not everything 
colloquially labeled as \emph{anthropomorphism} really is anthropomorphism.
\footnote{For instance, it is not always clear whether a person (may it be an adult or 
a child) who gives and anthropomorphic response, believes that the other agent really 
thinks or acts like a human or whether she is using anthropomorphism in a metaphoric 
sense.\citep{leeds_childrens_1992}} We need to be clear about what we mean by 
anthropomorphism because at the moment, there are different, even partially conflicting 
understandings of anthropomorphism, not only across but also within disciplines 
\cite{duffy_anthropomorphism_2002}. In particular, in the multi-disciplinary field of 
HRI, anthropomorphism is understood and interpreted in such a variety of ways that it 
is almost impossible to bring together and compare the research work done. In turn, it 
seems that the understanding of anthropomorphism in HRI has not improved much over the last decade but has only become more diverse (which is ok, too, but certainly a deeper 
understanding and common grounds would be convenient).

How can we narrow down the understanding of anthropomorphism? From a psychology point of view \cite{epley_when_2008} try to clarify anthropomorphism by explaining what it is \textit{not}, and we think this approach could also help researchers in HRI and robotics. \cite{epley_when_2008} list four effects that anthropomorphism does not include:
(1) First, anthropomorphism does not include behavioral descriptions of observable actions. Anthropomorphism requires going beyond what is directly observable.
(2) Second, anthropomorphism does not merely entail animism, as animate life is not a uniquely human-like characteristic.
(3) Third, anthropomorphism does not include any requirement of reasoned or reflective endorsement of an inference.\footnote{\hl{(the strength of anthropomorphic inferences can vary)}}
(4) Finally, anthropomorphism is not necessarily inaccurate. A psychological theory should take this (in)accuracy in anthropomorphism into account and predict variability in the tendency to perceive human-like traits in non-human agents.\footnote{For more details, the reader may refer to the original work \citep{epley_when_2008}.}

Not only \cite{epley_when_2008} and \cite{duffy_anthropomorphism_2002} but also others have noticed that anthropomorphism is a very complex phenomenon -- itself and in how it can be understood. This raises question how to study and operationalize anthropomorphism in an appropriate way (see Section \ref{sec:measuring}}.
A common formal ground is missing so far; however, it is important to be clear about what exactly we mean when using the term anthropomorphism.

\hl{(maybe here Persson: anthropomorphism as a multi-layered phenomenon)}

Similarly \cite{ruijten_introducing_2014} argues that anthropomorphism

@@

The phenomenon of attributing animacy and intentions to nonliving objects such as simple shapes has been intensively studied in developmental psychology. As mentioned earlier, Piaget found that children tend to ascribe life to the nonliving. They also tend to interpret physical phenomena in terms of intention on the part of the non-human subject. For instance children are likely to argue that \textit{``the sun is hot because it wants to make people warm''} \citep{leeds_childrens_1992}. In this example it is in fact questionable whether children really anthropomorphize the sun (believing that it has an intention) or whether they are using anthropomorphism in a metaphoric sense (because their conceptions of the world and of the living are not yet fully formed). Later however, two famous experiments by \cite{heider_experimental_1944} and \cite{michotte_perception_1963} showed that also adult participants attribute animacy and intention to simple things and shapes based on motion.\footnote{For more details, the reader may refer to the ``attribution theory''.} In both experiments, participants viewed animations of simple shapes, such as circles or triangles (see Figure \ref{fig:animacy_attribution}). Asked to describe what they observed, most people developed elaborate stories, attributing motivations, emotions and relationships between the objects.

\begin{figure}[h]\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.6]{heider-simmel_animation.jpeg}
% figure caption is below the figure
 \caption{Fritz Heider and Mary-Ann Simmel's animated film to study people's tendency to attribute animacy and intention to simple shapes; \textit{An experimental study of apparent behavior}, 1944}
 \fxnote{wrap text around this figure?}
 \label{fig:animacy_attribution}       % Give a unique label
 \end{figure}

@@

To sum it up, while anthropomorphism in robotics is a commonly discussed and
studied trait of human-robot interaction, we think it paradoxically lacks a
formal ground and unified understanding. The understanding of anthropomorphism
in robotics seems to not take into account the wide range of phenomena that it
encompasses. The traditional understanding of anthropomorphism in robotics to
date considers two main factors that account for the social phenomenon: the
characteristics of \textbf{1) the robot's design} (degree of human-likeness) and
the psychological determinants in \textbf{2) the human user} (see
\cite{epley_seeing_2007}). It has been shown that both these factors can
facilitate or hinder anthropomorphism. We propose to go beyond this traditional
perception of anthropomorphism as a static feature of HRI: Based on Persson
\textit{et al.}'s~\cite{persson_anthropomorphism_2000} argumentation that
anthropomorphism is a multi-layered phenomenon which arises in different levels,
we suggest to understand anthropomorphism as a \emph{dynamic},
\emph{context-dependent} process. To this end, we apply Persson \textit{et
al.}'s six levels on anthropomorphism to HRI and introduce three
\emph{interaction phases}. These interaction phases of anthropomorphism reflect
also the results of a long-term study that we conducted in a real human
environment and provide an account of the so-called \textit{novelty effect}. We
also draw attention to two new determinants of anthropomorphism, namely, the
context of use and the fact of getting used to a system over time. We propose to
add these two aspects as a third factor, the characteristics of \textbf{3) the
interaction} (time and space), to the previous two factors of anthropomorphism
(robot and user).

@@

In this article we will use the term \emph{anthropomorphic effect} to denote
observable correlates of anthropomorphism. With these we include, amongst
others, verbal utterances (such as direct speech or the use of pronouns
"he/she"), (social) gestures (such as a hug, waving, deictic gestures) or
\hl{anticipation of certain actions of an agent [what??]}.

Anthropomorphism can also be viewed as an \textit{experience} that arises in the
\textit{interaction} between a set of user expectations and external reality
\cite{persson_anthropomorphism_2000},
(Figure~\ref{fig:anthropomorphism_and_interaction}). Persson \textit{et al.}
suggest that the social phenomenon involves several levels, like primitive
psychology, folk-psychology, social stereotypes, and emotional anthropomorphism.

\begin{figure}[ht!]\centering
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[scale=0.42]{persson_anthropomorphism.png}
% figure caption is below the figure
 \caption{Anthropomorphism emerges in the (real or imagined) interaction between
 robot and user; illustration taken from \cite{persson_anthropomorphism_2000}
 \hl{maybe we don't put the figure here, it was just to illustrate what I mean}}

 \label{fig:anthropomorphism_and_interaction}       % Give a unique label
 \end{figure}

We would like to take up the proposition to discriminate different levels of
anthropomorphism, since each has its own characteristics and involves specific
types of user expectations. This differentiation could allow to draw inferences
on the human-robot relationship from the observed level of anthropomorphism.

\begin{figure}
    \centering
    \includegraphics[width=0.6\columnwidth]{factors}
    \caption{}
    \label{factors}
\end{figure}

The notation \Ant[R,H,C,t] reflects these multiple factors: we suggest here that
anthropomorphism is a function of a vector of robot's traits $R = \{r_0,...,r_n\}$, a vector human
determinants $H = \{h_0,...,h_n\}$, a set of interaction situations that build a
an interaction context $C = \{s_0,...,s_n\}$, and the duration of the
interaction itself, $t$.

@@

>>>>

We openly explored long-term interactions between human users and a domestic
service robot in a 6-months lasting ethnographic study~\citep{fink_living_2013}. 
We gave a commercially
available vacuum cleaning robot (iRobot's Roomba, see Figure
\ref{fig:aibo-roomba-ipad} middle) to nine households. 
%in the French speaking
%part of Switzerland. 
%The fairly diverse sample of households consisted of 2
%single homes, 1 couple, and 6 families with children in their home. In total,
%this summed up to 30 participants: 15 adults (26-71 years old, mean age 43.6
%years), and 15 children (6 months to 18 years old). During the 6 months, each
%household was visited 5 times to conduct interviews and make field observations
%in the households. Moreover, during several weeks, households were asked to fill
%in a diary to report the usage of the robot and other cleaning activities.
Hereafter, we refer to this study as the ``Roomba study''. It is one of very few 
long-term human-robot interaction studies in the domestic environment. 
In the following, we present findings in respect of two aspects: First, we
qualitatively describe what kinds of anthropomorphic interactions between
participants and the robot were observed and outline how these interactions
evolved over time.  Second, we present people's perception of human-likeness and
social capabilities of the robot and how this perception changed over
time.

%\footnote{\hl{note that these two different kinds of data
%(behavior/interaction and perception) already show that we try to measure
%anthropomorphism from these two angles!}}

@@

The direction of causality in the artifact-centered view is that the non-human artifact 
(or a certain feature of it) encourages the human to anthropomorphize the artifact / 
agent. This perspective assumes that humans directly respond to life-like and social 
cues that the non-human agent or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it, and in turn apply 
human-human social schemas and norms to the occurring interactions.
This view is acceptable since humans naturally react to specific 
cues. From early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
or to respond to social cues, such as facial expressions, social gestures or turn-taking. 
One has to keep in mind that these cues are of various senses and modalities in human-
human interaction, and consequently human-like design also addresses several senses. 
One of the most outstanding aspects is probably the outer appearance of an agent or 
artifact (in particular when it is unknown or unfamiliar). \cite{schmitz_concepts_2011} 
argues that an agent's outer appearance is likely to have an important impact on how 
the object is overall perceived. Consequently, if a robot's outer form resembles a 
human, the robot is likely to be treated similar to a human (\eg eye-contact is 
established or the robot is greeted).

@@

\subsubsection{Anthropomorphic form can be advantageous for HRI\\}

@@

\subsubsection{Anthropomorphic form can be disadvantageous for HRI\\}

@@

Literature on anthropomorphism is quite diverse. There exist different understandings of what anthropomorphism is. 
One reason for this might be that anthropomorphism
is a concept that is studied in very different domains that might each
integrate diverse or even contradictory understandings, or simply refer to different things.
However, the conflict is, that despite the fact that there is no commonly accepted definition, the terms
\textit{anthropomorphism}, \textit{anthropomorphic} or \textit{human-like} are
often used as if their meanings were clear and agreed upon
\citep{persson_anthropomorphism_2000}. Duffy \cite{duffy_anthropomorphism_2002} and \cite{epley_when_2008}
argue that these terms might even be misused. For instance, some researchers refer to \textit{"the
robot's level of anthropomorphism"} \cite{bartneck_is_2007}, whereas others
disagree to such a statement because they argue that the system or artifact itself does not \textit{"contain anthropomorphism" per se} but only gives rise to the process of anthropomorphizing in a
given user and situation. Consequently, \cite{persson_anthropomorphism_2000} conclude that
anthropomorphism emerges in the \textit{interaction} between the technology and
the user, and has to be understood as such. 
We will come back to this view later, in the summary of this section (Subsection \ref{sec:multi-factors}).

Another paradox is that, while human's tendency to anthropomorphize is commonly observed, it is 
at the same time still rather poorly understood
\cite{epley_seeing_2007}. The phenomenon itself has been found to be very complex though sometimes
subtle and as such hard to study \cite{epley_when_2008,duffy_anthropomorphism_2002}. Further, it is not clear how to operationalize anthropomorphism. We would like to discuss the challenges of studying and measuring anthropomorphism toward the end of this article, in Section \ref{sec:measuring}.

%Anthropomorphism is used in different senses throughout different disciplines 
%\citep{duffy_anthropomorphism_2003} which makes it difficult to synthesize findings. 
Depending on which understanding of anthropomorphism is applied, not everything 
colloquially labeled as \emph{anthropomorphism} really is anthropomorphism.
One practical example is that, it is not always clear whether a person (may it be an adult or 
a child) who gives and anthropomorphic response, believes that the other agent really 
thinks or acts like a human or whether she is using anthropomorphism in a metaphoric 
sense.\citep{leeds_childrens_1992}.
Thus, we need to be clear about what we mean by anthropomorphism.

But how to clarify the understanding of anthropomorphism?
From a psychology point of view \cite{epley_when_2008} try to clarify anthropomorphism by explaining what it is \textit{not}, and we think this approach could also help researchers in HRI and robotics. \cite{epley_when_2008} list four effects that anthropomorphism does not include:
(1) First, anthropomorphism does not include behavioral descriptions of observable actions. Anthropomorphism requires going beyond what is directly observable.
(2) Second, anthropomorphism does not merely entail animism, as animate life is not a uniquely human-like characteristic.
(3) Third, anthropomorphism does not include any requirement of reasoned or reflective endorsement of an inference.\footnote{\hl{(the strength of anthropomorphic inferences can vary)}}
(4) Finally, anthropomorphism is not necessarily inaccurate. A psychological theory should take this (in)accuracy in anthropomorphism into account and predict variability in the tendency to perceive human-like traits in non-human agents.\footnote{For more details, the reader may refer to the original work \citep{epley_when_2008}.}

Similarly to the last point mentioned by \cite{epley_when_2008}, several researchers argue that there might not just be one single type of anthropomorphism but different levels or gradations. For instance, \cite{persson_anthropomorphism_2000} proposes to understand anthropomorphism as a multi-layered phenomenon, and suggests that it exists of different levels, namely primitive categorization, primitive psychology, folk-psychology, traits, social roles, and emotional anthropomorphism. Correspondingly, \cite{ruijten_introducing_2014} propose to understand anthropomorphism as a continuum ranging from low to high human likeness, and to measure it as such.

@@

The phenomenon of attributing animacy and intentions to nonliving objects such as simple shapes has been intensively studied in developmental psychology. As mentioned earlier, Piaget found that children tend to ascribe life to the nonliving. They also tend to interpret physical phenomena in terms of intention on the part of the non-human subject. For instance children are likely to argue that \textit{``the sun is hot because it wants to make people warm''} \citep{leeds_childrens_1992}. In this example it is in fact questionable whether children really anthropomorphize the sun (believing that it has an intention) or whether they are using anthropomorphism in a metaphoric sense (because their conceptions of the world and of the living are not yet fully formed). Later however, two famous experiments by \cite{heider_experimental_1944} and \cite{michotte_perception_1963} showed that also adult participants attribute animacy and intention to simple things and shapes based on motion.\footnote{For more details, the reader may refer to the ``attribution theory''.} In both experiments, participants viewed animations of simple shapes, such as circles or triangles 
%(see Figure \ref{fig:animacy_attribution})
. Asked to describe what they observed, most people developed elaborate stories, attributing motivations, emotions and relationships between the objects.

%\begin{figure}[h]\centering
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[scale=0.6]{heider-simmel_animation.jpeg}
%% figure caption is below the figure
% \caption{Fritz Heider and Mary-Ann Simmel's animated film to study people's tendency to attribute animacy and intention to simple shapes; \textit{An experimental study of apparent behavior}, 1944}
% \fxnote{wrap text around this figure?}
% \label{fig:animacy_attribution}       % Give a unique label
% \end{figure}

@@

To sum it up, while anthropomorphism is a commonly discussed and
studied trait of human-robot interaction, its the understanding 
does not seem to take into account the wide range of phenomena that it
encompasses. To date, the understanding of anthropomorphism in robotics considers  two main factors: the
characteristics of \textbf{1) the robot's design} (degree of human-likeness) and
the psychological determinants in \textbf{2) the human user} (see
\cite{epley_seeing_2007}). A lot of research has shown that both these factors can
facilitate or hinder anthropomorphism.

However, based on what we have just presented, we would like to make a note, that there are more than these two factors that impact on anthropomorphism.

Based on the understanding that anthropomorphism represents a specific type of \textit{experience} that arises in the
\textit{interaction} between a set of user expectations and external reality
\cite{persson_anthropomorphism_2000}, 
%(see Figure~\ref{fig:anthropomorphism_and_interaction})
we propose that
anthropomorphism is a multi-layered phenomenon which arises in different levels, and is likely to evolve with continued interaction.
%we suggest to understand anthropomorphism as a \emph{dynamic},
%\emph{context-dependent} process. To this end, we apply Persson \textit{et
%al.}'s six levels on anthropomorphism to HRI and introduce three
%\emph{interaction phases}. These interaction phases of anthropomorphism reflect
%also the results of a long-term study that we conducted in a real human
%environment and provide an account of the so-called \textit{novelty effect}.

To this end, we make the following two suggestions:

First, we propose to include two more main factors as determinants of anthropomorphism (besides the robot's design, and the human user's characteristics). Namely, these factors are \textbf{3) the context of use} (situation) and the fact of getting used to a system the more and longer one interacts with it, thus becoming familiar with the system over \textbf{4) time}.

%\begin{figure}[ht!]\centering
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[scale=0.42]{persson_anthropomorphism.png}
%% figure caption is below the figure
% \caption{Anthropomorphism emerges in the (real or imagined) interaction between
% robot and user; illustration taken from \cite{persson_anthropomorphism_2000}
% \hl{maybe we don't put the figure here, it was just to illustrate what I mean}}
%
% \label{fig:anthropomorphism_and_interaction}       % Give a unique label
% \end{figure}

Second, we would like to take up the proposition already made by other researchers (\eg \cite{persson_anthropomorphism_2000,ruijten_introducing_2014}) to not viewing anthropomorphism as being either there or not there, but to discriminate different levels of
anthropomorphism (low, and high anthropomorphism, for instance).
This seems to make sense, since each of the four proposed factors has its own characteristics and involves specific
types of user expectations, for instance. Such a differentiation in the understanding of anthropomorphism could further allow to draw inferences
on the human-robot relationship based on the observed level of anthropomorphism.

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{factors}
    \caption{Anthropomorphism is a multi-factor phenomenon. How far people tend to anthropomorphize a robot is determined by four main factors: the design of the robot $(R)$, the personal characteristics of the human user $(H)$, the situation and context in which the interaction occurs $(C)$, and the amount of time (repeated interaction) that the user spends to familiarize herself with the robot $(t)$.}
    \label{factors}
\end{figure}

To facilitate things, we introduce the notation \Ant[R,H,C,t] to reflect these multiple factors: we suggest here that
anthropomorphism is a function of a vector of robot's traits $R = \{r_0,...,r_n\}$, a vector human
determinants $H = \{h_0,...,h_n\}$, a set of interaction situations that build
an interaction context $C = \{s_0,...,s_n\}$, and the duration of the
interaction itself, $t$.

@@

%In this article we will use the term \emph{anthropomorphic effect} to denote
%observable correlates of anthropomorphism. With these we include, amongst
%others, verbal utterances (such as direct speech or the use of pronouns
%"he/she"), (social) gestures (such as a hug, waving, deictic gestures) or
%\hl{anticipation of certain actions of an agent [what??]}.

% Persson \textit{et al.}
%suggest that the social phenomenon involves several levels, like primitive
%psychology, folk-psychology, social stereotypes, and emotional anthropomorphism.

@@

Also, the methods and measurements applied to study anthropomorphism appear insufficient, when applying a holistic understanding of anthropomorphism emerging from an interaction between a human user and a robot, for instance. Our point here is that to study a phenomenon as complex as anthropomorphism, we believe that one needs to take into account many factors (see Section \ref{sec:multi-factors}), besides looking at the human-like design of the artifact.
However, most of the
experiments that investigate anthropomorphism in robotics are conducted in
controlled lab-settings during short-term human-robot interactions. This
experimental setting can be critical for studying
anthropomorphism and findings might possibly lead to interpretations that would not be generalizable to natural settings or real interaction scenarios. For instance, some studies are about showing pictures of different robots to human subjects and then asking them how "human-like" they would rate these robots. What points of reference do the participants have in this setting? What can you really tell from a picture? I found that some normal people totally misinterpret the size / height of a robot, when it is not indicated. Also, when showing only pictures, movements, behaviors, materials, and so forth are totally disregarded. However, these aspects are also crucial to assess how "human-like" a robot appears. Last but not least, it has been found that "human-likeness" is a quite diverse and abstract term that includes many things. Instead of asking such a big question, \cite{kahn_jr._robotic_2006} suggest to ask for more concrete constructs that are typical or unique of the concept of "human-likeness", such as being hungry or feeling guilty. For instance, \cite{ruijten_introducing_2014} propose a 25 concrete questions to measure anthropomorphism along a scale that accounts for lower and higher levels of it.
Some studies (e.g. exploring characteristics of the "uncanny valley") also seem to contrast two or
more different anthropomorphic or mechanical systems that might not be
appropriate for comparison given the complex context.
